Modeling Ionic Polymer-Metal Composites\\ with Space-Time Adaptive Multimesh \emph{hp}-FEM

### 2010-10-25 results.tex
Minor fixes
New iteration, Pavel, October 24.

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=.8\columnwidth]{mesh}
  \caption{\label{fig:mesh} Initial coarse mesh (a),
  	half refined mesh (b) and refined mesh (c). The coarse mesh (a)
	and refined mesh (c) were used in the initial calculations, the latter one
	in case of \emph{p}-adaptivity (including HP\_ANISO\_P). The half-refined mesh was
	used later to optimize the \emph{hp}-adaptive refinement solutions.}
  \end{centering}
\end{figure}
Calculation for all the refinement modes were performed
in both single-mesh and multi-mesh configurations. 
The following numerical results were recorded for each 
refinement mode: converged relative error, cumulative CPU
time, and the problem size in terms of number
of degrees of freedoms (NDOFs) at each time step. 
Two types of initial meshes were used --- in case of only \emph{p}-adaptivity,
more refined mesh was used (Fig.~\ref{fig:mesh}~(c)) to ensure
the error convergence.
When the element size refinement
was also enabled (all \emph{h}/\emph{hp} refinement modes), very coarse initial mesh
was used (see Fig.~\ref{fig:mesh}~(a)) to let the adaptivity
algorithm find the most optimal mesh. That coarse initial mesh might not be
suitable for practical applications, as will be demonstrated in the
end of this section, however, it provides a good insight into the
adaptivity performance of Hermes.
In both cases, the initial mesh was loaded at each time step of the
calculation.

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=.75\columnwidth]{cphi}
  \caption{\label{fig:cphi} Concentration $C$
  and voltage $\phi$ at two different time steps
  (HP\_ANISO refinement mode was used). This is a 2D solution shown
  in 3D, where the height indicates the values of $C$ and $\phi$.}
  \end{centering}
\end{figure}
Example of the solution at $t=0.7\ s$ and $t=3.0\ s$ 
calculated with HP\_ANISO refinement mode is shown
in Fig.~\ref{fig:cphi}. The time $t=0.7\ s$ was chosen because
by that time (with given constants in Table~\ref{Table:used-constants}),
some ionic migration has already taken place and
the concentration gradients near the boundaries $\partial\Omega_1$ and
$\partial\Omega_3$ have formed. The automatic mesh refinements
at different time steps are clearly visible in the figure --- especially
near $\partial\Omega_1$, where the $C_{\Omega_1}\approx 0$ and
the gradient $\nabla C$ is
shifting in time (this is also demonstrated in Fig.~\ref{fig:comsol-conc-volt}).

The following subsections provide a detailed comparison of
the different refinement modes and try to estimate the most suitable
one for the given problem.

\subsection{Optimal refinement modes}

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{singlemulti_dof}
  \caption{\label{fig:singlemultidof} NDOFs in case 
  of single-mesh and multi-mesh configurations with HP\_ISO
  and HP\_ANISO refinement modes. (Notice log Y scale)}
  \end{centering}
\end{figure}

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{singlemulti_cpu}
  \caption{\label{fig:singlemulticpu} Cumulative CPU times in case
  of single-mesh and multi-mesh configurations with HP\_ISO
  and HP\_ANISO refinement modes. (Notice log Y scale)}
  \end{centering}
\end{figure}

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=.75\columnwidth]{poly}
  \caption{\label{fig:poly} Polynomial degree space
  for $C$ and $\phi$ at $t=0.7\ s$ The color indicates
  the maximum polynomial degree of the corresponding element.}
  \end{centering}
\end{figure}

Running the simulation with different refinement modes 
and meshes showed that the multi-mesh configuration generally results in
a smaller problem, faster calculation, and better or similar error convergence
compared to the single-mesh configuration.
This is well illustrated in Fig.~\ref{fig:singlemultidof} 
and Fig.~\ref{fig:singlemulticpu}.
The result can be understood when considering Fig.~\ref{fig:cphi} --- in the vicinity
of the boundaries $\partial \Omega_1$ and $\partial\Omega_3$ the concentration gradient
is greate than voltage gradient, namely $\nabla C >> \nabla \phi$. Therefore,
adapting the mesh for both variables is not reasonable in terms of
number of degrees of freedom. For instance,
the solution space with corresponding mesh in case of
HP\_ANISO at $t=0.7\ s$ is shown in Fig.~\ref{fig:cphi}. The corresponding polynomial
degree space for multi-mesh configuration is shown in Fig.~\ref{fig:poly}. 
Notice that the adaptivity algorithm
of Hermes has increased the maximum polynomial degree for the $C$-space to~7, however,
the maximum polynomial degree for the $\phi$-space is~2. Furthermore,
the mesh is significantly more refined for $C$.
For most of the cases using the multi-mesh results in similar or better CPU time.
It means that multi-mesh takes less computational resources. Only excpetion was HP\_ANISO\_H 
refinement mode for which the single mesh configuration resulted slightly faster
calculation time. 
Based on the results, only multi-mesh configurations will be considered
in the following comparisons.

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoaniso_dof}
  \caption{\label{fig:isoanisodof} NDOFs in case 
  of multi-mesh configuration with H\_ISO, H\_ANISO,
  HP\_ISO, HP\_ANISO, and single-mesh configuration with HP\_ANISO\_H
  refinement modes. (Notice log Y scale)}
  \end{centering}
\end{figure}

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoaniso_cpu}
  \caption{\label{fig:isoanisocpu} Cumulative CPU times in case 
  of multi-mesh configuration with H\_ISO, H\_ANISO,
  HP\_ISO, HP\_ANISO, and single-mesh configuration with HP\_ANISO\_H
  refinement modes. (notice log Y scale)}
  \end{centering}
\end{figure}

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoanisop_dof}
  \caption{\label{fig:isoanisopdof} NDOFs in case 
  of multi-mesh configuration with P\_ISO, P\_ANISO, and
  HP\_ANISO\_P refinement modes.}
  \end{centering}
\end{figure}

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoanisop_cpu}
  \caption{\label{fig:isoanisopcpu} Cumulative CPU times in case 
  of multi-mesh configuration with P\_ISO, P\_ANISO, and
  HP\_ANISO\_P refinement modes.}
  \end{centering}
\end{figure}
To narrow down the list of refinement modes for the given problem, first the
isotropic and anisotropic adaptivities were compared. The \emph{p}-adaptivity
modes were compared separately as they used more refined initial mesh 
(shown in Fig.~\ref{fig:mesh}~(c)).
Fig.~\ref{fig:isoanisodof} and Fig.~\ref{fig:isoanisocpu} provide the comparison
of H\_ISO, H\_ANISO, HP\_ISO, HP\_ANISO, and HP\_ANISO\_H modes in terms of 
cumulative CPU time and NDOFs.
Fig.~\ref{fig:isoanisopdof} and Fig.~\ref{fig:isoanisopcpu} show the similar
comparsion for the P\_ISO, P\_ANISO, and HP\_ANISO\_P modes.
It is not difficult to see that the anisotropic refinement modes result in a reasonable problem
size and the problems solve within a reasonable calculation time. It is interesting
to note that HP\_ANISO results in the smallest problem size. Also, in
\emph{p}-adaptivity group, HP\_ANISO\_P results in the smallest problem size
at each time step, whereas P\_ISO and P\_ANISO result in a very large problem size
during the first time steps of the solution.
Here the term ``reasonable problem size''
means that the number of degrees of freedom in time converges
to so that $N_{DOF}<500$, and the term ``reasonable calculation time''
means that the calculation (step $\tau=0.01\ s$, physical
time $t_{end}=3.0\ s$) time $t$ on a givem system was $t<500\ s$.
Although these parameters are empirical, they serve as an upper limit, given
that the most refinement modes give significantly better results ---
$t<<500\ s$ and $N_{DOF} << 500$.

\subsection{Quantitative analysis of the refinement modes}

Based on the results in the previous subsection, H\_ANISO, HP\_ANISO,
and HP\_ANISO\_H from the \emph{hp/h}-adaptivity group and HP\_ANISO\_P
from \emph{p}-adaptivity group will be compared
in terms of NDOFs and cumulative CPU time. 
In all of the cases, the relative 
error at each time step remained below
the threshold which was set to $e_{th}=0.5\%$ between the coarse mesh
and fine mesh solutions, therefore the error-time plot will not be considered.

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{cpu}
  \caption{\label{fig:cpu} Comparative cumulative CPU time for different refinement modes.}
  \end{centering}
\end{figure}

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{dof}
  \caption{\label{fig:dof} Comparative NDOFs at each time step for 
  different refinement modes.}
  \end{centering}
\end{figure}

Fig.~\ref{fig:cpu} shows the cumulative CPU time for different refinement 
modes at each time step --- the results were recorded on the same computer
for the CPU times to be comparable.
We see that HP\_ANISO\_H and HP\_ANISO require the most
resources. This can be understood from the fact that in case of these 
refinement modes the adaptivity algorithm has the largest number of 
candidates (see Section.~\ref{sec:model} and Fig.~\ref{fig:refinements})
from which a particular refinement can be chosen.
At the same time, HP\_ANISO\_P is the fastest compared to the other refinement modes.
This can be attributed to the finer inital mesh.

Fig.~\ref{fig:dof} shows the NDOFs at each time step.
It can be seen that the HP\_ANISO results in the 
smallest problem size --- $N_{DOF} \approx 125$. 
All the other refinement modes result in a 
problem size of approximately ($N_{DOF} \approx 250$). Therefore
the HP\_ANISO is the most efficient in terms of memory consumption.

As we have filtered out
two notable refinement modes for the given problem --- HP\_ANISO\_P because of the fast
calculation, and HP\_ANISO because of the small problem size. In the following
subsection, some optimizations to reduce the CPU time of
HP\_ANISO and reduce the NDOFs of HP\_ANISO\_P will be considered.

\subsection{Optimizations of HP\_ANISO and HP\_ANISO\_P refinement modes}

>>>>

The solutions to the PNP problem exhibit a specific behavior that was 
described above. In order to find the best adaptive method to deal with 
this type of problems, we performed numerous computations using all 
adaptivity modes in both the single-mesh and multi-mesh regimes.
In the numerical experiments we paid attention to the 
relative error, cumulative CPU time, and problem size 
in terms of number of degrees of freedom (DOF) in each 
time step.

We used two types of initial meshes --- a finer mesh shown 
in Fig.~\ref{fig:mesh}~(c) was used for \emph{p}-adaptivity
and a very coarse initial mesh shown in  Fig.~\ref{fig:mesh}~(a) was 
used for $h$-adaptivity and $hp$-adaptivity.

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=.8\columnwidth]{mesh}
  \caption{\label{fig:mesh} Initial coarse mesh (a),
  	half refined mesh (b) and refined mesh (c). The coarse mesh (a)
	and refined mesh (c) were used in the initial calculations, the latter one
	in case of \emph{p}-adaptivity (including HP\_ANISO\_P). The half-refined mesh (b) was
	used later to optimize the \emph{hp}-adaptive refinement solutions.}
  \end{centering}
\end{figure}

An example of the solution at $t=0.7\ s$ and $t=3.0\ s$ 
calculated with the HP\_ANISO refinement mode is shown
in Figs.~\ref{fig:cphi-1} and \ref{fig:cphi-2}.

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=.75\columnwidth]{cphi-1}
  \caption{\label{fig:cphi-1} Concentration $C$
  and voltage $\phi$ at $t=0.7\ s$.}
  \end{centering}
\end{figure}

\newpage

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=.75\columnwidth]{cphi-2}
  \caption{\label{fig:cphi-2} Concentration $C$
  and voltage $\phi$ at $t=3.0\ s$.}
  \end{centering}
\end{figure}

The reader can see that at $t=0.7\ s$ some ionic migration has already 
taken place and large concentration gradients near the boundaries $\partial\Omega_1$ 
and $\partial\Omega_3$ have formed. The figures also show that the meshes 
at $t=0.7\ s$ and $t=3.0\ s$ are different.

\subsection{Comparison of single mesh low-order FEM and $hp$-FEM}

Comparison of single mesh H\_ANISO with p=1, single mesh H\_ANISO with p=2
and single mesh HP\_ANISO should be added here to make the point that 
hp-FEM is much better than low-order FEM. 
Otherwise the reader does not know 
why we are just using hp-FEM in the following. Or is there a reason 
why you do not want to add this here --  long computing time for hp-FEM or something 
else?

\subsection{Comparison of single-mesh and multi-mesh $hp$-FEM}

Running the simulation with different adaptivity modes 
and meshes showed that the multi-mesh $hp$-FEM configuration resulted into
the smallest problems, shortest computing times, and better or the same error 
convergence compared to any single-mesh configuration.
This is illustrated in Figs.~\ref{fig:singlemultidof} 
and Fig.~\ref{fig:singlemulticpu}.

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{singlemulti_dof}
  \caption{\label{fig:singlemultidof} Number of degrees of freedom (DOF) as a function 
  of physical time for single-mesh and multi-mesh configurations with the HP\_ISO
  and HP\_ANISO adaptivity modes. Note logarithmic scale on the vertical axis.}
\vspace{-6mm}
  \end{centering}
\end{figure}

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{singlemulti_cpu}
  \caption{\label{fig:singlemulticpu} Cumulative CPU time as a function 
  of physical time for single-mesh and multi-mesh configurations with the HP\_ISO
  and HP\_ANISO adaptivity modes. Note logarithmic scale on the vertical axis.}
  \end{centering}
\end{figure}

\noindent 
Fig. \ref{fig:poly} shows higher-order meshes in the adaptive multimesh $hp$-FEM
computation for $C$ and $\phi$ at $t = 0.7$ s. Different 
colors mean different polynomial degrees. A diagonal pattern inside an element 
tells that the element has different polynomial degrees in the 
horizontal and vertical directions.

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=.75\columnwidth]{poly}
  \caption{\label{fig:poly} Higher-order FEM mesh for 
  $C$ and $\phi$ at $t=0.7\ s$. }
  \end{centering}
\end{figure}

These result are in good agreement with Fig.~\ref{fig:cphi-2} -- in the vicinity
of the boundaries $\partial \Omega_1$ and $\partial\Omega_3$, the concentration gradient
is much greater than the voltage gradient. Therefore the multimesh $hp$-FEM adaptivity 
algorithm has increased the maximum polynomial degree for the $C$-space to~7 while 
the maximum polynomial degree for the $\phi$-space is~2. One can also see that the mesh 
is significantly more refined for $C$.
Since these results are representative for all adaptivity modes, only multi-mesh 
configurations are considered in the following.

\subsection{Comparison of isotropic and anisotropic refinements}

Next we would like to illustrate the role of anisotropic mesh refinements.
Figs.~\ref{fig:isoanisodof} and \ref{fig:isoanisocpu} show typical results 
for the H\_ISO, H\_ANISO, HP\_ISO, HP\_ANISO adaptivity modes in terms 
of DOF and cumulative CPU time.

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoaniso_dof}
  \caption{\label{fig:isoanisodof} Number of DOF as a function of physical time for 
  multi-mesh configurations with H\_ISO, H\_ANISO,
  HP\_ISO, and HP\_ANISO adaptivity modes.}
  \end{centering}
\end{figure}

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoaniso_cpu}
  \caption{\label{fig:isoanisocpu} Cumulative CPU time as a function of physical time 
  for multi-mesh configurations with H\_ISO, H\_ANISO,
  HP\_ISO, and HP\_ANISO adaptivity modes.}
  \end{centering}
\end{figure}

\noindent
Figs.~\ref{fig:isoanisopdof} and \ref{fig:isoanisopcpu} present a similar 
comparsion for the P\_ISO, P\_ANISO, and HP\_ANISO\_P modes. Recall that these 
computations use a different initial mesh that was a-priori refined in space.

\newpage

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoanisop_dof}
  \caption{\label{fig:isoanisopdof} Number of DOF as a function of physical time
  for multi-mesh configurations with P\_ISO, P\_ANISO, and
  HP\_ANISO\_P adaptivity modes.}
  \end{centering}
\end{figure}

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoanisop_cpu}
  \caption{\label{fig:isoanisopcpu} Cumulative CPU times as a function of physical time
  for multi-mesh configurations with P\_ISO, P\_ANISO, and
  HP\_ANISO\_P adaptivity modes.}
  \end{centering}
\end{figure}

As a conclusion, the reader can see that the anisotropic adaptivity modes always perform better than 
the isotropic ones. In particular, HP\_ANISO results into the smallest problem size. 
In the \emph{p}-adaptivity group, HP\_ANISO\_P leads to a small problem size
consistently in each time step, whereas P\_ISO and P\_ANISO yield large problems
during the first few time steps.

%
%\subsection{Quantitative analysis of the refinement modes}
%
%Based on the results in the previous subsection, H\_ANISO, HP\_ANISO,
%and HP\_ANISO\_H from the \emph{hp/h}-adaptivity group and HP\_ANISO\_P
%from \emph{p}-adaptivity group will be compared
%in terms of NDOFs and cumulative CPU time. 
%In all of the cases, the relative 
%error at each time step remained below
%the threshold which was set to $e_{th}=0.5\%$ between the coarse mesh
%and fine mesh solutions, therefore the error-time plot will not be considered.
%
%\begin{figure}[!ht]
%  \begin{centering}
%  \includegraphics[width=\columnwidth]{cpu}
%  \caption{\label{fig:cpu} Comparative cumulative CPU time for different refinement modes.}
%  \end{centering}
%\end{figure}
%
%\begin{figure}[!ht]
%  \begin{centering}
%  \includegraphics[width=\columnwidth]{dof}
%  \caption{\label{fig:dof} Comparative NDOFs at each time step for 
%  different refinement modes.}
%  \end{centering}
%\end{figure}
%
%Fig.~\ref{fig:cpu} shows the cumulative CPU time for different refinement 
%modes at each time step --- the results were recorded on the same computer
%for the CPU times to be comparable.
%We see that HP\_ANISO\_H and HP\_ANISO require the most
%resources. This can be understood from the fact that in case of these 
%refinement modes the adaptivity algorithm has the largest number of 
%candidates (see Section.~\ref{sec:model} and Fig.~\ref{fig:refinements})
%from which a particular refinement can be chosen.
%At the same time, HP\_ANISO\_P is the fastest compared to the other refinement modes.
%This can be attributed to the finer inital mesh.
%
%Fig.~\ref{fig:dof} shows the NDOFs at each time step.
%It can be seen that the HP\_ANISO results in the 
%smallest problem size --- $N_{DOF} \approx 125$. 
%All the other refinement modes result in a 
%problem size of approximately ($N_{DOF} \approx 250$). Therefore
%the HP\_ANISO is the most efficient in terms of memory consumption.
%
%As we have filtered out
%two notable refinement modes for the given problem --- HP\_ANISO\_P because of the fast
%calculation, and HP\_ANISO because of the small problem size. In the following
%subsection, some optimizations to reduce the CPU time of
%HP\_ANISO and reduce the NDOFs of HP\_ANISO\_P will be considered.

\subsection{Optimization of the HP\_ANISO and HP\_ANISO\_P adaptivity modes}
